---
editor_options: 
  chunk_output_type: console
---

**OH BLOODY HELL THERE'S TIDYVERSE EVERYWHERE REMOVE IT ALL**

# Getting data

This section focusses on accessing and downloading WATLAS data. This is done using functions in the [WATLAS Utilities package](https://github.com/pratikunterwegs/watlastools). 

**Workflow**

1. Preparing required libraries.
2. Reading tag data with deployment start dates from a local file. This file is not yet publicly available.
3. Connecting to the NIOZ databse and downloading data. This database is also not public-access.

## Prepare `watlastools` and other libraries

```{r install_watlastools, message=FALSE, warning=FALSE}
# install the package watlastools from master branch using the following
# install.packages("devtools")
library(devtools)

devtools::install_github("pratikunterwegs/watlastools")
library(watlastools)

# libraries to process data
library(data.table)
library(ggplot2)
library(ggthemes)
library(purrr)
library(glue)
```

## Read in tag deployment data

```{r get_deployment_data, message=FALSE, warning=FALSE}
# read deployment data from local file in data folder
tag_info <- fread("data/data2018/SelinDB.csv")

# filter out NAs in release date and time
tag_info <- tag_info[!is.na(Release_Date) & !is.na(Release_Time),]

# make release date column as POSIXct
tag_info[,Release_Date := as.POSIXct(paste(Release_Date, 
                                           Release_Time, sep = " "), 
                                     format = "%d.%m.%y %H:%M", tz = "CET")]

# check new release date column
head(tag_info$Release_Date)
```

```{r plot_release_schedule, echo=FALSE, fig.cap="Knots released per week of 2018.", message=FALSE, warning=FALSE}
# check release cohort
ggplot(tag_info)+
  geom_bar(aes(x = week(Release_Date)), col = 1, fill = "grey")+
  ggthemes::theme_few()+
  labs(x = "release week (2018)", y = "# knots released",
       caption = Sys.time())
```

## Get data and save locally

```{r get_acess, warning=FALSE, message=FALSE}
# read in database access parameters from a local file
data_access <- fread("data/access_params.txt")
```

```{r get_data, message=FALSE, warning=FALSE}
# create a data storage file if not present
# use the getData function from watlastools on the tag_info data frame
# this is placed inside a pmap wrapper to automate access for all birds

if(!dir.exists("data/data2018")) { dir.create("data/data2018") }

pmap(tag_info[,.(Toa_Tag, Release_Date)], function(Toa_Tag, Release_Date){

  prelim_data <- watlastools::wat_get_data(tag = Toa_Tag,
                           tracking_time_start = as.character(Release_Date),
                           tracking_time_end = "2018-10-31",
                           username = data_access$username,
                           password = data_access$password)
  
  setDT(prelim_data)
  # prelim_data[,TAG:= = as.numeric(TAG) - prefix_num]

  message(glue('tag {Toa_Tag} accessed with {nrow(prelim_data)} fixes'))
  
  fwrite(prelim_data, file = glue('data/data2018/{Toa_Tag}_data.csv'),
         dateTimeAs = "ISO")
  
})
```

