---
editor_options: 
  chunk_output_type: console
---

# Adding tidal cycle data

This section is about adding tidal cycle data to individual trajectories. This is done to split the data up into convenient, and biologically sensible units. This section uses the package `VulnToolkit` [@VulnToolkit2014] to identify high tide times from water-level data provided by Rijkswaterstaat for the measuring point at West Terschelling.

**Workflow**

1. Prepare required libraries,
2. Read in water level data and identify high tides,
3. Write tidal cycle data to local file,
4. Add time since high tide to movement data.

## Prepare libraries

```{r install_vulntoolkit_2, message=FALSE, warning=FALSE}
# load VulnToolkit or install if not available
if ("VulnToolkit" %in% installed.packages() == FALSE) {
  devtools::install_github("troyhill/VulnToolkit")
}
library(VulnToolkit)

# libraries to process data
library(data.table)
library(purrr)
library(glue)
library(dplyr)
library(stringr)
library(fasttime)
library(lubridate)

library(watlastools)
```

## Read water level data

Water level data for [West Terschelling](https://waterinfo.rws.nl/#!/details/publiek/waterhoogte-t-o-v-nap/West-Terschelling(WTER)/Waterhoogte___20Oppervlaktewater___20t.o.v.___20Normaal___20Amsterdams___20Peil___20in___20cm), a settlement approx. 10km from the field site are provided by Rijkswaterstaat's [Waterinfo](https://waterinfo.rws.nl/#!/nav/bulkdownload/parameters/Waterhoogten/), in cm above Amsterdam Ordnance Datum. These data are manually downloaded in the range July 1, 2018 -- October 31, 2018 and saved in `data/data2018`.

```{r read_waterlevel_data, message=FALSE, warning=FALSE}
# read in waterlevel data
waterlevel <- fread("data/data2018/waterlevelWestTerschelling.csv", sep = ";")

# select useful columns and rename
waterlevel <- waterlevel[, .(WAARNEMINGDATUM, WAARNEMINGTIJD, NUMERIEKEWAARDE)]

setnames(waterlevel, c("date", "time", "level"))

# make a single POSIXct column of datetime
waterlevel[, dateTime := as.POSIXct(paste(date, time, sep = " "),
  format = "%d-%m-%Y %H:%M:%S", tz = "CET"
)]

waterlevel <- setDT(distinct(setDF(waterlevel), dateTime, .keep_all = TRUE))
```

## Calculate high tides

A tidal period of 12 hours 25 minutes is taken from [Rijkswaterstaat](https://www.rijkswaterstaat.nl/water/waterdata-en-waterberichtgeving/waterdata/getij/index.aspx).

```{r get_high_tide, message=FALSE, warning=FALSE}
# use the HL function from vulnToolkit to get high tides
tides <- VulnToolkit::HL(waterlevel$level,
  waterlevel$dateTime,
  period = 12.41,
  tides = "H", semidiurnal = TRUE
)

# read in release data and get first release - 24 hrs
tag_info <- fread("data/data2018/SelinDB.csv")
tag_info <- tag_info[!is.na(Release_Date) & !is.na(Release_Time), ]
tag_info[, Release_Date := as.POSIXct(paste(Release_Date,
  Release_Time,
  sep = " "
),
format = "%d.%m.%y %H:%M", tz = "CET"
)]

# remove NAs
tag_info <- na.omit(tag_info, cols = "Release_Date")

first_release <- min(tag_info$Release_Date, na.rm = TRUE) - (3600 * 24)

# remove tides before first release
tides <- setDT(tides)[time > first_release, ][, tide2 := NULL]
tides[, tide_number := 1:nrow(tides)]
```

```{r write_tide_data, message=FALSE, warning=FALSE}
# write to local file
fwrite(tides,
  file = "data/data2018/tides_2018.csv",
  dateTimeAs = "ISO"
)
```

## Add time since high tide

```{r time_to_high_tide, message=FALSE, warning=FALSE}
# read in data and add time since high tide
data_files <- list.files(
  path = "data/data2018/data_processed/",
  pattern = "id_", full.names = TRUE
)
data_ids <- str_extract(data_files, "(id_\\d+)") %>% str_sub(-3, -1)

# map read in and tidal time calculation over data
# merge data to insert high tides within movement data
# arrange by time to position high tides correctly
invisible(
  lapply(data_files, function(df) {

    # read and fix data types
    temp_data <- fread(df, integer64 = "numeric")
    temp_data[, ts := fastPOSIXct(ts, tz = "CET")]

    # merge with tides and order on time
    temp_data <- wat_add_tide(
      data = temp_data,
      tide_data = "data/data2018/tides_2018.csv"
    )

    # add waterlevel
    temp_data[, temp_time := lubridate::round_date(ts, unit = "10 minute")]
    temp_data <- merge(temp_data, waterlevel[, .(dateTime, level)],
      by.x = "temp_time", by.y = "dateTime"
    )
    setnames(temp_data, old = "level", new = "waterlevel")

    # export data, print msg, remove data
    fwrite(temp_data, file = df, dateTimeAs = "ISO")
    message(glue("tag {unique(temp_data$id)} added time since high tide"))
    rm(temp_data)
  })
)
```

