---
editor_options: 
  chunk_output_type: console
---

# Residence patch cleaning

Residence patches must be cleaned to remove patches where the speed between patches is too high, or too low. We filter for $0 < v < 150$, where $v$ is the speed between patches.

We also filter out patches that are too small, ie, fewer than 10 fixes (approx. 5 minutes).

Once cleaned, these patches can be passed on to the overlap finding algorithm.

## Prepare libraries

```{r prep_libs, message=FALSE, warning=FALSE}
library(data.table)
library(purrr)
```

## Load patches

```{r}
# load data
patches <- fread("data/data2018/data_2018_patch_summary.csv")

# add uid
patches[, uid := seq_len(nrow(patches))]
```

## Calculate speed between patches

```{r}
# this is the speed in metres per second
patches[, speed := distBwPatch / 
          c(NA, time_start[-1] - 
              time_end[seq_len(length(time_end) - 1)]),
        by = .(id, tide_number)]

# what is 150 km/hr in m/s
cutoff_speed <- 15 # around 54 kmph

# filter ridiculous speeds
patches <- patches[between(speed, 1e-4, cutoff_speed), ]
```

## Remove sequences of reflected patches

```{r}
# recalculate the distance b/w patches
patches2 <- patches[id == "554", ]
patches2 <- patches2[duration >= 10 * 60, ]

patches2 <- split(patches2, by = c("id"))


# this is the speed in metres per second
patches2 <- map(patches2, function(df) {
  
  df <- df[speed < 15, ]
  
  df[, distBwPatch := watlastools::wat_bw_patch_dist(df)]
  
  # replace 
  df[, distBwPatch := nafill(distBwPatch, type = "const", fill = Inf)]
  
  df[, speed := distBwPatch / 
          c(1, time_start[-1] - 
              time_end[seq_len(length(time_end) - 1)])]
  
  # where is the speed greater than a cutoff
  df[, reflection := cumsum(speed > cutoff_speed)]
  
  return(df)
})
```

```{r}
# bind the rows
patches2 <- rbindlist(patches2)
patches2[tide_number == 96, ]
```


The number of patches is down to 57,089.

## Export this 'good' patch data

```{r}
# export the summary
fwrite(patches, "data/data2018/data_2018_good_patches.csv")
```
