---
editor_options: 
  chunk_output_type: console
---

# Residence patch construction

This section is about using the main `watlastools` functions to infer residence points when data is missing from a movement track, to classify points into residence or travelling, and to construct low-tide residence patches from the residence points. Summary statistics on these spatial outputs are then exported to file for further use.

**Workflow**

1. Prepare `watlastools` and required libraries,
2. Read data, infer residence, classify points, construct low-tide patches, and get spatial data from patches,
3. Get patch trajectories,
4. Export spatial data.

## Prepare libraries

```{r prep_libs, message=FALSE, warning=FALSE, eval=FALSE}
# load watlastools or install if not available
if("watlastools" %in% installed.packages() == FALSE){
  devtools::install_github("pratikunterwegs/watlastools")
}
library(watlastools)

# libraries to process data
library(dplyr)
library(data.table)
library(purrr)
library(stringr)
library(glue)
library(readr)
library(fasttime)

# functions for this stage alone
ci <- function(x){qnorm(0.975)*sd(x, na.rm = T)/sqrt((length(x)))}
```

## Patch construction

```{r remove_old_data, eval=FALSE}
if(file.exists("data/data2018/data_2018_tidal_mean_speed.csv")){
  file.remove("data/data2018/data_2018_tidal_mean_speed.csv")
}

if(file.exists("data/data2018/data_2018_patch_summary.csv")){
  file.remove("data/data2018/data_2018_patch_summary.csv")
}
```

Process patches. Takes approx. 5 hours for 3 second data.

```{r make_patches, eval=FALSE, message=FALSE, warning=FALSE}
# make a vector of data files to read
data_files <- list.files(path = "data/data2018/revisitData", 
                         pattern = "_revisit.csv", full.names = TRUE)

# get tag ids
data_id <- str_split(data_files, "/") %>% 
  map(function(l)l[[4]] %>% str_sub(1,3)) %>% 
  flatten_chr()

# make df of tag ids and files
data <- data_frame(tag = data_id, data_file = data_files)
data <- split(x = data, f = data$tag) %>% 
  map(function(l) l$data_file)

# map inferResidence, classifyPath, and getPatches over data
for(i in 1:length(data)){
  patch_data <- map(data[[i]], function(l){
  
    # read the data file
    temp_data <- fread(l)
    temp_data[,ts:=fastPOSIXct(ts)]
    
    id <- unique(temp_data$id)
    tide_number <- unique(temp_data$tide_number)
    
    # get data summary for the tidal cycle
    {
      data_summary <- temp_data[,.(duration = (max(time) - min(time))/60,
                                   n_fixes = length(x),
                                   prop_fixes = length(x) / ((max(time) - min(time))/30)),
                                by = .(id, tide_number)]
      
      sld <- watlastools::wat_simple_dist(temp_data, "x", "y")
      timelag <- c(NA, as.numeric(diff(temp_data$time)))
      speed <- sld/timelag
      
      data_summary[,`:=`(mean_speed = mean(speed, na.rm=TRUE),
                         ci95_speed = ci(speed))]
      
      # write total distance for the tidal cycle
      fwrite(data_summary, file = "data/data2018/data_2018_tidal_mean_speed.csv", 
             append = TRUE, scipen = 6)
    }
    
    # wrap process in try catch
    tryCatch(
      {
        # watlastools function to infer residence
        temp_data <- wat_infer_residence(df = temp_data,
                                         infPatchTimeDiff = 30,
                                         infPatchSpatDiff = 100)
        
        # watlastools function to classify path
        temp_data <- wat_classify_points(somedata = temp_data,
                                         resTimeLimit = 2)
        
        # watlastools function to get patches
        patch_dt <- wat_make_res_patch(somedata = temp_data,
                                         bufferSize = 10,
                                         spatIndepLim = 100,
                                         tempIndepLim = 30,
                                         restIndepLim = 30,
                                         minFixes = 3)
        # print message
        message(as.character(glue('patches {id}_{tide_number} done')))
        return(patch_dt)
      },
      # null error function, with option to collect data on errors
      error= function(e)
      {
        message(glue::glue('patches {id}_{tide_number} errored'))
      }
    )
  })
  
  # repair high tide patches across an individual's tidal cycles
  repaired_data <- wat_repair_ht_patches(patch_data_list = patch_data)
  # write patch summary data
  
  if(all(is.data.frame(repaired_data), nrow(repaired_data) > 0)){
    # watlastools function to get patch data as spatial
    patch_summary <- wat_get_patch_summary(resPatchData = repaired_data,
                                           whichData = "summary")
    fwrite(patch_summary, file = "data/data2018/data_2018_patch_summary.csv",
           append = TRUE)
  }
}

```
