---
editor_options: 
  chunk_output_type: console
---

# Finding networks in tidal cycles

## Read in patch data and overlaps

```{r}
# to handle data
library(readr)
library(scales)
library(tidyr); library(tibble)
library(magrittr)
library(dplyr)
library(purrr)

# to work with networks
library(igraph)
```

```{r}
# read in patch data
data_patches <- read_csv("data/data2018/data_2018_patch_summary_has_patches.csv")
# read in overlap data
data <- read_csv("data/data2018/data_spatio_temporal_overlap_2018.csv")
```

## Determine tidal cycle and tidal stage

### Link overlaps with patch data

```{r}
# add a uid
data_patches <- mutate(data_patches, uid = 1:nrow(data_patches))

# select columns
data_patches <- select(data_patches,
                       id, uid, tide_number, waterlevel_mean)

# link overlaps with patches
data <- left_join(data, data_patches,
                  by = c("patch_i_unique_id" = "uid")) %>% 
  left_join(data_patches,
            by = c("patch_j_unique_id" = "uid"))
```

### Get consensus tide number waterlevel

The tide number and tidal stage will be used to divide the overlaps into classes, so that we can examine community structure in a particular tide and at a certain tidal stage.

```{r}
# get the consensus tide as first and the mean waterlevel
data <- data %>% 
  mutate(grp_id = 1:nrow(.)) %>% 
  group_by(grp_id) %>% 
  mutate(tide_number = tide_number.x,
         waterlevel = mean(c(waterlevel_mean.x, waterlevel_mean.y))) %>% 
  ungroup()

# remove bad cols and classify as high or low tide
data <- data %>% 
  ungroup() %>% 
  select(-grp_id) %>% 
  rename(id_x = id.x, id_y = id.y) %>% 
  select(-contains(c(".x", ".y")))

# classify as high or low
data <- data %>% 
  mutate(tide_stage = if_else(waterlevel < 55, "low", "high")) %>% 
  select(-waterlevel)
```

## Prepare network data

### Prepare subset-wise edgelist

Here, the subsets are the tide number and tidal stage. The edge-list indicates pairwise overlap.

```{r}
# get the total overlap between pairs as the edges df
edges_df <- data %>% 
  group_by(id_x, id_y, tide_number, tide_stage) %>% 
  summarise_at(vars(temporal_overlap_seconds, spatial_overlap_area),
               .funs = sum) %>% 
  filter(id_x != id_y)

# group and nest
edges_df <- group_by(edges_df, tide_number, tide_stage) %>% 
  nest()
```

### Prepare subset-wise node list

```{r}
# make nodes data -- this the individual identities
# add individual data to patch data
data_id <- read_csv("data/data2018/SelinDB.csv") %>% 
  filter(!is.na(Toa_Tag)) %>% 
  select(Toa_Tag, 
         wing = WING, mass = MASS, 
         gizzard_mass) %>% 
  distinct(Toa_Tag, .keep_all = TRUE) %>%  
  rename(id = Toa_Tag)

# expand nodes df to match edges data
nodes_df <-  
  summarise(edges_df, 
            id = map(data, function(df) {union(df$id_x, df$id_y)})) %>%
  mutate(id = map(id, function(df){
    
    df <- tibble(id = df) %>% 
      left_join(data_id, by = "id")
  }))
```

This data needs to be converted to `igraph` networks.

## Network metrics with igraph

### Convert to igraph

```{r}
# prepare a uniform network dataframe
network_df <- left_join(edges_df, nodes_df)

# prepare network objects
network_df <- mutate(network_df,
                     net = map2(data, id, function(e, n){
                       tmp_net <- igraph::graph_from_data_frame(d = e,
                                                                directed = FALSE,
                                                                vertices = n)
                     }))
```

### WIP
