---
editor_options: 
  chunk_output_type: console
---

# Finding networks in tidal cycles

## Read in patch data and overlaps

```{r}
# to handle data
library(readr)
library(scales)
library(tidyr); library(tibble)
library(magrittr)
library(dplyr)
library(purrr)

# to work with networks
library(igraph)

# to plot
library(ggplot2)
library(scico)
```

```{r}
# read in patch data
data_patches <- read_csv("data/data2018/data_2018_patch_summary_has_patches.csv") %>% 
  mutate(uid = as.character(1:nrow(data_patches)))
# read in overlap data
data <- read_csv("data/data2018/data_spatio_temporal_overlap_2018.csv")
```

## Identify communities of residence patches

### Make unweighted network from overlaps

```{r}
# prepare edgelist
data_edges <- select(data, contains("patch"))

# prepare network from edgelist
network_overlap <- igraph::graph_from_data_frame(d = data_edges, directed = F)
```

### Find communities

```{r}
# get communities using fastgreedy
# the network is unweighted
data_communities <- igraph::fastgreedy.community(network_overlap)
d <- communities(data_communities) %>% 
  imap(function(.x, .id){
    tibble(flock = .id,
           patch = .x)
  })

# get data as a tibble
flock_id <- bind_rows(d)
```

### Join flock data to edgelist

```{r}
# join flock and patch by uid on the first patch
data <- data %>% 
  mutate_at(vars(matches("patch")),
            as.character) %>% 
  left_join(flock_id,
            by = c("patch_i_unique_id" = "patch"))

# join patch data, basically id data, to flock data
tmp_patches <- filter(data_patches,
                      uid %in% flock_id$patch) %>% 
  select(id, uid)
data <- left_join(data, tmp_patches,
                  by = c("patch_i_unique_id" = "uid")) %>% 
  left_join(tmp_patches,
            by = c("patch_j_unique_id" = "uid"))

# get id and weights
edges_df <- select(data,
               matches("(id.)|overlap|flock")) %>% 
  rename(id_x = id.x, id_y = id.y)
```

## Prepare network data

### Prepare subset-wise edgelist

Here, the subsets are the tide number and tidal stage. The edge-list indicates pairwise overlap. We now add subsets for 500m grid cells on the landscape.

```{r}
# get the total overlap between pairs as the edges df
edges_df <- edges_df %>% 
  group_by(flock, id_x, id_y) %>% 
  summarise_at(vars(temporal_overlap_seconds, spatial_overlap_area),
               .funs = sum) %>% 
  filter(id_x != id_y)

# group and nest
edges_df <- group_by(edges_df, 
                     flock) %>% 
  nest()

# count flock edges and remove small flocks
# ie, with fewer than 3 edges
edges_df <- mutate(edges_df,
                   n_edges = map_int(data, nrow)) %>% 
  filter(n_edges > 3)
```

### Get individual data

```{r}
# make nodes data -- this the individual identities
# add individual data to patch data
data_id <- readxl::read_excel("data/data2018/Biometrics_2018-2019.xlsx") %>% 
  filter(str_detect(`TAG NR`, "[a-zA-Z]+", negate = TRUE))

# a function for gizzard mass
get_gizzard_mass <- function(x, y) {-1.09 + (3.78*(x*y))}

# add gizzard mass
data_id <- mutate(data_id,
                  gizzard_mass = get_gizzard_mass(SH1, SW1))

# rename columns and drop ids without mass and gizzard mass
data_id <- data_id %>% 
  select(id = `TAG NR`, 
         wing = WING, mass = MASS, 
         gizzard_mass) %>% 
  distinct(id, .keep_all = TRUE) %>% 
  drop_na(mass, gizzard_mass)

# add some exploration scores and tag info
data_behav <- read_csv("data/data2018/2018-19-all_exploration_scores.csv") %>% 
  filter(Exp == "F01")
data_tag <- read_csv("data/data2018/tag_info.csv") %>% 
  mutate(id = as.character(Toa_Tag))

# join all scores
data_id <- left_join(data_id, data_tag,
                     by = c("id")) %>% 
  left_join(data_behav, by = "FB")

# remove ids with no exploration
data_id <- mutate(data_id,
                  behav = Mean) %>% 
  # drop_na(behav) %>% 
  select(id, mass, gizzard_mass, behav)
```

### Prepare nodes data

```{r}
# expand nodes df to match edges data
nodes_df <-  
  summarise(edges_df, 
            id = map(data, 
                     function(df) {union(df$id_x, df$id_y)})) %>%
  mutate(id = map(id, function(df){
    
    df <- tibble(id = as.character(df)) %>% 
      left_join(data_id, by = "id")
    
    # remove nodes with no attributes
    df <- drop_na(df, mass, gizzard_mass, behav)
  }))
```

This data needs to be converted to `igraph` networks.

## Network metrics with igraph

### Convert to igraph

```{r}
# prepare a uniform network dataframe
network_df <- left_join(edges_df, nodes_df)

# filter for existing node data
network_df <- filter(network_df,
                     map_lgl(id, function(df){nrow(df)>1}))

# prepare network objects for assortnet via igraph
network_df <- mutate(network_df,
                     net = map2(data, id, function(e, n){
                       
                       # exclude edges with no vertex attributes
                       e <- filter(e, 
                                   id_x %in% n$id,
                                   id_y %in% n$id)
                       
                       tmp_net <- igraph::graph_from_data_frame(d = e,
                                                                directed = FALSE,
                                                                vertices = n)
                     }))
```

### Get assortativity

Remove networks with few (< 3) edges.

```{r}
# check which graphs are okay
network_df <- mutate(network_df, 
                     valid_graph = map_int(net, function(n){
                       E(n) %>% length()
                     })) %>% 
  ungroup()

# remove bad graphs, ie, with fewer than 3 edges
# this is both for quality and practical reasons
min_edges <- 3
network_df <- filter(network_df, valid_graph > min_edges)
```


```{r}
# get weighred assortativity by gizzard mass
network_df <- mutate(network_df,
                     assort_g_mass = unlist(purrr::map(net, function(n){
                       
                       adj <- igraph::as_adjacency_matrix(n,
                                                          sparse = FALSE,
                                                          attr = "spatial_overlap_area")
                       
                       assortnet::assortment.continuous(adj, V(n)$gizzard_mass)
                     })))
```

```{r}
# get weighred assortativity by gizzard mass
network_df <- mutate(network_df,
                     assort_behav = unlist(purrr::map(net, function(n){
                       
                       adj <- igraph::as_adjacency_matrix(n,
                                                          sparse = FALSE,
                                                          attr = "spatial_overlap_area")
                       
                       assortnet::assortment.continuous(adj, V(n)$behav)
                     })))
```

### Plot histogram of assortativity

```{r}
# prepare the data
plot_df <- select(network_df, 
                  flock, valid_graph,
                  contains("assort")) %>% 
  filter(valid_graph >= 3) %>% 
  pivot_longer(cols = contains("assort"))
```

```{r eval=FALSE}
ggplot(plot_df)+
  geom_histogram(aes(x = value),
                 col = "black", size =0.3,
                 fill = "grey", 
                 binwidth = 0.2)+
  geom_vline(xintercept = 0, lty = 2, colour = "red")+
  facet_grid(~name)
```

## Visualise assortativity over space

### Link flock assortativity and patch data

```{r}
# subset network data
data_sp_assort <- select(network_df,
                         flock, valid_graph,
                         contains("assort"))

# add flock data to patch data
data_patches <- left_join(data_patches,
                          flock_id,
                          by = c("uid" = "patch"))

# get mean x, y and waterlevel for flocks
data_flock_attrs <- select(data_patches,
                           uid,
                           matches("(x|y|waterlevel)(_mean)"),
                           flock) %>% 
  filter(uid %in% flock_id$patch) %>% 
  group_by(flock) %>% 
  summarise_at(vars(contains("mean")),
               .funs = mean)

# join flock attributes to assort data and classify tide
data_sp_assort <- left_join(data_sp_assort,
                            data_flock_attrs) %>% 
  mutate(tide_stage = if_else(waterlevel_mean < 55, "low", "high"))

# melt data
data_sp_assort <- select(data_sp_assort,
                         -waterlevel_mean) %>% 
  pivot_longer(cols = contains("assort"))
```

### Plot assort in space

```{r}
# get griend
griend <- sf::st_read("griend_polygon/")

# examine differences high and low tide
ggplot(data_sp_assort)+
  geom_jitter(aes(x_mean,y_mean, col = value),
            alpha = 0.5, size = 2)+
  geom_sf(data = griend, fill = NA, size = 0.5)+
  # geom_histogram(aes(assort_g_mass,
  #                fill = tide_stage),
  #              position = "identity",
  #              alpha = 0.5)+
  # coord_cartesian(ylim = c(-0.5, 0.5))+
  scale_colour_gradient2(low = "indianred",
                       mid = "white",
                       high = "blue")+
  # coord_sf(crs = 32631)+
  facet_grid(name~tide_stage, scales = "fixed")+
  theme_grey()+
  theme(legend.position = "top",
        legend.key.height = unit(0.2, "cm"),
        panel.border = element_rect(fill=NA,
                                        colour = "black"),
        axis.text = element_blank(),
        axis.title = element_blank())+
  coord_sf(expand = F)

ggsave("figs/fig_assort_gizzard.png", dpi = 600)
```

