---
editor_options: 
  chunk_output_type: console
---

# Linking patch metrics to personality

This section is about linking patch metrics to individual personality measures, using the previously derived patch summary data, and experimentally measured exploration scores as a proxy for personality.

**Workflow**

1. Prepare required libraries,
2. Link patch metrics with individual behaviour score,
3. Run models for patch metrics of interest,
4. Plot figures for metrics ~ personality,
5. Export figures and model output.

## Prepare libraries

```{r prep_libs_2, message=FALSE, warning=FALSE, eval=FALSE}
# libraries to process data
library(readr) # using readr here since data is small
library(purrr)
library(glue)
library(dplyr)
library(tidyr)

# libraries for stats
library(lmerTest)

# libraries for plotting
library(ggplot2)
library(ggthemes)
library(scico)
library(scales)
library(gridExtra)
library(cowplot)

# ci func
# simple ci function
ci <- function(x){
  qnorm(0.975)*sd(x, na.rm = T)/sqrt((length(x)))}
```

## Read in exploration score

```{r link_patches_explorescore, eval=FALSE, message=FALSE, warning=FALSE}
# read in tag info
tag_info <- read_csv("data/data2018/SelinDB.csv") %>% 
  select(RINGNR, Toa_Tag) %>% 
  filter(!is.na(Toa_Tag))

# read in updated behavioural data
behav_score <- read_delim("data/data2018/Selindb-updated_2019-07-17.csv", delim = ";")

behav_score <- inner_join(tag_info, behav_score, by = c("RINGNR" = "FB")) %>% 
  filter(trial == "F01") %>% 
  select(Toa_Tag, texpl)
```

## Exploratory analysis on patch metrics

```{r eda_patches, eval=FALSE, message=FALSE}
# read in patch data
patch_data <- read_csv("data/data2018/patch_summary.csv")

# process data to plot histograms of area, duration, circularity, displacement
# propfixes and tidal time
patch_data <- patch_data %>% 
  select(id, tide_number, area, circularity, duration, dispInPatch, propfixes,
         tidaltime_mean, distBwPatch) %>% 
  pivot_longer(cols = c(area, circularity, duration, dispInPatch, propfixes,
                        tidaltime_mean, distBwPatch),
               names_to = "variable",
               values_to = "value")

# filter out 95% values in each variable
patch_data <- patch_data %>% 
  group_by(variable) %>% 
  mutate(scaled_val = rescale(value)) %>% 
  filter(scaled_val < 0.90)

# plot histograms
ggplot(patch_data)+
  geom_histogram(aes(x = value))+
  facet_wrap(~variable, scales = "free")+
  theme_bw()

```


## Patch metrics as an outcome of exploration score

```{r patch_models, eval=FALSE, message=FALSE, warning=FALSE}
# read in patch data
patch_data <- read_csv("data/data2018/patch_summary.csv")
# filter for patch quality on the proportion of expected fixes acheived
patch_data <- filter(patch_data, propfixes %between% c(0.2, 1.0))

# join to behav data
patch_data <- inner_join(patch_data, behav_score, by=c("id" = "Toa_Tag")) %>% 
  mutate(texpl = as.numeric(texpl)) %>% 
  filter(!is.na(texpl))

# prepare for simultaneous modelling
model_data <- setDF(patch_data) %>% 
  pivot_longer(names_to = "response_variable", values_to = "response_value",
               cols = c(duration, dispInPatch, distBwPatch, area)) %>%
  group_by(response_variable) %>%
  nest()

# run models using glmerTest and get predict output
model_data <- mutate(model_data, 
                     model = map(data, function(z){
                       lmer(response_value ~ texpl + 
                              (1|tide_number) + 
                              tidaltime_mean + I((tidaltime_mean)^2), 
                            data = z, na.action = na.omit)})) %>% 
  
  mutate(predMod = map2(model, data, function(a, b){
    mutate(b, predval = predict(a, type = "response", newdata = b, 
                                allow.new.levels = TRUE))
  }))

# assign names to model object
names(model_data$model) <- glue("response = {model_data$response_variable} | predictor = tExplScore")
```

## Plot model figures

```{r plot_model_figs, eval=FALSE, message=FALSE}
# prepare data for plotting
plot_data <- mutate(model_data,
                    data_expl = map(data, function(df){
                      select(df, texpl, response_value, tide_number) %>% 
                        filter(tide_number < 100) %>% 
                        mutate(texpl = plyr::round_any(texpl, 0.1)) %>% 
                        group_by(texpl) %>% 
                        summarise_at(vars(response_value), 
                                     .funs = c(~mean(., na.rm=T), ~ci(.)))}))
                    
# prepare plots to arrange
plots <- plot_data %>% 
  ungroup() %>% 
  mutate(y = c("time in patch (mins)",
               "disp in patch (m)",
               "dist b/w patch (m)",
               "patch area (m.sq.)"),
         title = glue::glue('{letters[1:4]} {y}')) %>% 
  
  # plots column
  mutate(plot = pmap(select(., data_expl, y, title), function(data_expl, y, title){
    ggplot(data_expl)+
      geom_pointrange(aes(x = texpl,
                          y = mean,
                          ymin = mean-ci,
                          ymax = mean+ci),
                      position = position_dodge(width = 0.01))+
      scale_y_continuous(labels = scales::comma)+
#      facet_grid(~phase, scales = "free_x")+
      #scico::scale_colour_scico(palette = "cork")+
      
# 
#       coord_cartesian(xlim = c(0,1),
#                       ylim = c(quantile(data_expl$mean, 0.00, na.rm = TRUE),
#                                quantile(data_expl$mean, 0.999, na.rm = TRUE)))+

      scale_x_continuous(breaks = c(0,0.5,1))+
      theme_few()+
      theme(legend.position = "none",
            #plot.background = element_rect(colour = 1),
            #plot.margin = unit(rep(0.5, 4), "cm"),
            #axis.text.y = element_text(angle = 90, vjust = 0.5),
            plot.title = element_text(face = "bold"))+
      labs(y = y, title = title,
           x = "exploration score")
  }),
  plot = map(plot, cowplot::as_grob))

# arrange plots

plotlist = plots$plot

{
  # pdf(file = "../figs/fig05patchMetrics.pdf", width = 8, height = 12)
  grid.arrange(grobs = plotlist, ncol = 2)
  # dev.off()
}
```
                    
