---
output: html_document
editor_options: 
  chunk_output_type: console
---

# modified for red knots by prg p.r.gupte@rug.nl ####

this code generates data for testing the method of social network randomizations the code generates a population of agents with given movement rules. currently it is set on the paired agents scenario, see details below for switching between different scenarios or exploring the parameter spcae fo other values

writen by Orr Spiegel Jan 2016. contact me at orr.spiegel@mail.huji.ac.il if you need further help

```{r}
#Required packages #######
rm(list=ls())
require(CircStats); require(spatstat); require(sp); require(R.matlab);
library(RColorBrewer);library(scico)
```

```{r set model parameters}
#### model parameters and protocol selection for users to choose #######
ToPlot=1

#can be 0 (no plot) or 1 (create plots)- note this will slow the proccess a lot!!

Color_indv=c(scico(40, palette = "lapaz"))

#if ToPlot=01 this determines how many individfuals to plot (currently 12). The colors for plotted individuas: 2-=red 3=green 5= light green 6 purpule 7 yelow 8 grey

DayLength=59;

#how many time step for each day (currently it is limited to 59 since i considered it as minutes within an hour)

#'prg: 1 minute positions as here

DaysToSimulate=6 ;

#how many days to simulate

N_indv=40

#number of individuals in the population

Social_Pecrt_rng= 50

#detection range (in meters) - indivduals within this range will be considered in the bias point in the relevan scenario of sociable agents. set to 0 if you want socially indiferent agents (or in the fixed pairs scenario).

#'prg: a social bias/observation range of 5m seems realistic for short shorebirds moving within a larger group; individuals beyond this range are likely undetected by obstructed view or focusing distance.
#'prg: q: what is the focusing distance of knots? what is their visual system like? a cursory reading, possibly from other waders.

DriftHRCenters=1

#1 or 0 for a drift of the center in space along the simulation (i.e. the changing environment scenarios), 0 is no drift.

#'prg: there are two model options, one beginning at tidal maximum (high or low tide) and the other of t hours either side of tidal maximum (ie, TM - t --> TM + t). in the former, the centre of the movement track will shift towards positions at the next tidal maximum. these positions could be forced into the RW as an end point, or left free, and would develop over consecutive runs. it would interesting to see whether the final show NND similar to the empirical data, which would help resolve if the empirical pattern is from tidal forcing (from LT --> HT) and tidal release (HT --> LT) or from association.


DriftStrength=c(0,1000)

#if DriftHRCenters==1 this defines drift in m per day for X and Y values in the simulation 

#'prg: the movement centre is expected to drift at some m/day. over the mean tidal period of 6 hours, the drift would be d/4 metres. this distance should be calculated from the empirical data as the displacement between positions at tidal maxima. knots 1 data could be used as well, being more extensive 

DriftasOnce=2;

#if DriftHRCenters==1 this defines when does the drift occur. can be 0,1,2. if 0 it will drift daily; if 1 it will drift in the midle of the run if 2 it will drift from begining.

#'prg: drift must begin at the head of the run.

PropDriftIndiv=1;

#proportion of drifting individuals.

#'prg: all inividuals must drift

PairedAgents=0; 

#if 1 then simulates the paired agent scenario where agents are paired in their intial location and show attraction to their pair only, indiferent to the others.

#'prg: agents are not paired

PairStartDist=200;

#if PairedAgents=1 this sets the Upper limit of the distance of the pairs. For the second individual in each pair the location will be random within this range from the first mate.

#'set seed for randomisations
set.seed(1234) 

#comment it for new randomizations start points


#### agents movement rules: #### 
EtaCRW=0.7 

#the weight of the CRW component in the BCRW used to model the Indiv movement

#'prg: why 0.7? should this come from the data?

StpSize_ind=6 

#Mean step lengths of individuals;

#'prg: in what units? the map? in which case it should be 1cm/s, the same speed as the tide. with a simulation running at 10 minute intervals, that's 6m per step, assuming all indiviudals are identical.


StpStd_ind=5 

# Sandard deviations of step lengths of individuals

#'prg: what is a reasonable limit? 5 appears too high, let's make it 1.5

Kappa_ind=3 

#concentration parameters of von Mises directional distributions used for individuals' movement

#'prg: what is a reasonable limit? not clear, leave it as is
```

```{r plot_step_size_dist}
# plotting step size distribution: #'prg: plot option is off

if (ToPlot==1) {plot(density(rgamma(10000, shape = StpSize_ind^2/StpStd_ind^2,scale = StpStd_ind^2/StpSize_ind)))  
    rose.diag(rvm(n=10000, mean = 0, k = Kappa_ind),bins=72, pch = 16, cex = 1, shrink = 1)  }
```


```{r prep_dfs}
#### prepraration of variables for storing data ####

if (PairedAgents==1){Social_Pecrt_rng=0}

#if i work in a paired design than no social perception range for other agents

N_tmStp=DaysToSimulate*DayLength 

#  total namber of time steps that witll be simulated for each iteration

N_Rows=N_indv*N_tmStp

#'prg: set the number of rows in the df
```

```{r make_dfs}
#'prg: create a dataframe to hold the eventual values?

XYind_log2=as.data.frame(t(rbind(
  rep(seq(1:N_indv),length.out=N_Rows,each=N_tmStp),
  #indiviuals within iteration
  rep(seq(1:N_tmStp),length.out=N_Rows),
  #cumul time steps for each indiv
  rep(seq(1:DaysToSimulate),length.out=N_Rows,each=DayLength),
  #day in the simulation
  rep(seq(1:DayLength),length.out=N_Rows,each=1),
  #step within a day
  rep(seq(1:(DaysToSimulate*N_indv)),length.out=N_Rows,each=DayLength),
  #day in the simulation#burst (i.e., each day as a unique value, not same for two indiv)
  matrix(data=NA,nrow=3,ncol=N_Rows))));#state, x, y pseudoSex


#'assign names to the df and make some factors

names(XYind_log2)=c("indiv","step","Day","StepInDay","burst","x","y","pseudoSex")
XYind_log2$indiv=as.factor(XYind_log2$indiv);
XYind_log2$burst=as.factor(XYind_log2$burst);
```

```{r set_home_range_params}
#'set start index

startIndx=1;
Phi_ind <- rep(0, N_indv) 

#dierction of the last step for the individuals

XYind <- vector("list")

#this list will store the matrixes of locations of all individual

HRCnt=matrix(data=NA,nrow=N_indv,ncol=3)

#'this is related to the paired individuals scenario
#HRCnt[,3]=rep(c(1,2),N_indv/2) #since sex is listed in this parameter i update it to be pairs of diferent sexes for the paired design
#HRcenterDist=rep(0,N_indv/2);#Distance between pairs of agents at the begining
```

```{r home_range_drift}
#### setting drift of HR center (bias point) for the drifting scenarios ####
if (DriftasOnce==0){ #0 it will drift daily;
  Xdrift=cumsum(rep(c(DriftStrength[1],rep(0,(DayLength-1))),DaysToSimulate))#for each day i have the drift, then steps with no drift ubtil the next day. i work with Cumsum since the drift grows 
  Ydrift=cumsum(rep(c(DriftStrength[2],rep(0,(DayLength-1))),DaysToSimulate)) 
  print('using drift daily')
} else if (DriftasOnce == 1){ # if 1 it will drift in the midle of the run  
  Xdrift=rep(0,N_tmStp);Xdrift[N_tmStp/2]=DriftStrength[1]*DaysToSimulate;Xdrift=cumsum(Xdrift);
  Ydrift=rep(0,N_tmStp);Ydrift[N_tmStp/2]=DriftStrength[2]*DaysToSimulate;Ydrift=cumsum(Ydrift);  
  print('using drift from middle step of the run')
} else if (DriftasOnce == 2){#if 2 it will drift from begining.
  Xdrift=rep(DriftStrength[1]*DaysToSimulate,N_tmStp);
  Ydrift=rep(DriftStrength[2]*DaysToSimulate,N_tmStp);
  #print('using drift from start')
}else{print('check your params')}

DrifByStep=rbind(Xdrift,Ydrift)
```

```{r which_indivs_drift}
#Determining indivduals to drift.
DriftingYorN=c(rep(1,PropDriftIndiv*N_indv),rep(-1,N_indv-PropDriftIndiv*N_indv))
DriftingYorN=sample(DriftingYorN, N_indv);

#setting the HR centerDrift for this time step (cumulative from beggining)
if (DriftHRCenters==1){CurDrift=DrifByStep[,1];names(CurDrift)=NULL}else{CurDrift=c(0,0)}
```

```{r generate_landscape}
#### generating landscape for this iteration #######      
N_FoodPatches=50 #how many patches - this is just to be consistent with the code. currently individuals do not response to resources. 
Scl=7e3;#the size of the area for simuation. not important since there is not boundry. just to be consistent with other codes.
PointsStrc <- rsyst(nx=sqrt(N_FoodPatches))#generation of a uniform landscape
FoodItems=SpatialPoints(coords(PointsStrc)*Scl-Scl/2) # These are the points on scale of 0 to 1 no -Scl/2 to Scl/2
if (ToPlot==1) {plot(FoodItems,col='white',pch=1); title('the simulated landscape')}#the number of point will be more or less kappa*mu    
```

```{r set_initial_conditions}
#### a loop on individuals for initial conditions ####
for (k in 1:N_indv) {
  #placing agents in their initial positions 
  XYind[[k]] <- matrix(rep(NA, 2 * N_tmStp), ncol = 2)#this matrix will store the location of each individual
  XYind[[k]][1, ] <-  c(runif(n=1,min=-Scl/3,max=Scl/3 ),#X random initial location of each individual is uniformly distributed in the range of the observed ones
                         runif(n=1,min=-Scl/3,max=Scl/3 ))#Y random initial location of each individual is uniformly distributed in the range of the observed ones
  HRCnt[k,1:2] =XYind[[k]][1, ];#logging the HR center 
   
  #if ((PairedAgents==1) & (k %% 2 == 0)){#if this is a paired scenario than i switch every other individual to be a pair with nearby location to the previous one
   #  XYind[[k]][1, ]=XYind[[k-1]][1, ]+c(runif(n=2,min=-PairStartDist/2,max=PairStartDist/2))#setting the start tlocation for the pair
   #  HRcenterDist[k/2]=round(dist(rbind(XYind[[k-1]][1, ],c(XYind[[k]][1, ]))),digit=1);#storing the initial distance from this pair
  #   print(paste("pair:",k-1,"and",k, "distance",HRcenterDist[k/2] ))
#   }#if paired design than use nearby coordinate for every second individual

 #  if (ToPlot==1 ) {
 #     plot(SpatialPoints(coords=matrix(data=XYind[[k]][1, ],nrow=1)),add=T,pch=HRCnt[k,c(3)]+23,col=HRCnt[k,c(3)],bg=HRCnt[k,c(3)], cex=1.5) 
 #     if ((PairedAgents==1) & (k %% 2 == 0)){#plot also line between pairs
 #       plot(SpatialLines(list( Lines( list(Line(rbind(XYind[[k-1]][1, ],c(XYind[[k]][1, ])))), ID=k/2) )),add=T) # This is the river 'line' 
 #       }#plot also line between pairs
 #   }#plot the line
}#loop on individuals 
```


```{r run_positions}
#### loop on time steps and individuals to run the simulation ####
for (Curr_tmStp in 1:(N_tmStp-1)){
  ## loop on individuals ##
  for (Curr_indv in 1:N_indv){
          
  #### distance of this Curr_indv to all other individuals 
  Dist=rep(NA,N_indv); 
  for (ii in 1:N_indv) {Dist[ii] = dist(rbind(XYind[[ii]][Curr_tmStp, ],c(XYind[[Curr_indv]][Curr_tmStp, ])))}
  Dist[Dist==0]=NA;#getting rid of the distance to self
          
  ## start generation of indiv location at time k+1
          
  ##### selecting direction ##########
  #calculating the direction to the initial location (bias point )+ now with drift for the current step
  BiasPoint=(XYind[[Curr_indv]][1, ]+CurDrift*DriftingYorN[Curr_indv])#this bias point is the origin+ the current cumulative bias
 # if ((PairedAgents==1) & (Curr_indv %% 2 == 1)){BiasPoint=colMeans(rbind(BiasPoint,XYind[[Curr_indv+1]][Curr_tmStp, ]))}#updating bias point as the mean of HR center and mates last location
#  if ((PairedAgents==1) & (Curr_indv %% 2 == 0)){BiasPoint=colMeans(rbind(BiasPoint,XYind[[Curr_indv-1]][Curr_tmStp-1, ]))}#updating bias point as the mean of HR center and mates previous location (before last step since he moved already)  
          
  if (min(Dist,na.rm=T)< Social_Pecrt_rng) {#if there is another invividual in range, 
    #if i want the mean of direction to the HR and nearest neighbor: 
    #BiasPoint=colMeans(rbind(BiasPoint,XYind[[which.min(Dist)]][Curr_tmStp, ]))}#then updated  this for the bias also
    BiasPoint=XYind[[which.min(Dist)]][Curr_tmStp, ]
    }#then updated  this for the bias also
          
          
  coo <- BiasPoint  - XYind[[Curr_indv]][Curr_tmStp, ] #cheking direction to the Bias point of this individual, change the second ccomp to have another bias
  mu <- Arg(coo[1] + (0+1i) * coo[2])#
  if (mu < 0)  {   mu <- mu + 2 * pi  }#just making sure direction is not negative
  mu.av <- Arg(EtaCRW * exp(Phi_ind[Curr_indv] * (0+1i)) + (1 - EtaCRW) * exp(mu * (0+1i))) #bias to initial location + CRW to find the von mises center for next step
  Phi_ind[Curr_indv] <- rvm(n=1, mean = mu.av, k = Kappa_ind)#choosing curr step direction from vonMises centered around the direction selected above 
          
          
##### preforming step #########
#selection of step size for this indiv in this state from the specific gamma          
  step.len <- rgamma(1, shape = StpSize_ind^2/StpStd_ind^2,scale = StpStd_ind^2/StpSize_ind)
  step <- step.len * c(Re(exp((0+1i) * Phi_ind[Curr_indv])), Im(exp((0+1i) * Phi_ind[Curr_indv])))
  XYind[[Curr_indv]][Curr_tmStp + 1, ] <- XYind[[Curr_indv]][Curr_tmStp, ] + step#the indiv next location

###### ploting the steps ########
if (Curr_indv<=length(Color_indv) & ToPlot==1 ){#plot only up to 6 indiv and only if ToPlot==1
    StepAsLine = SpatialLines(list( Lines( list(Line(XYind[[Curr_indv]][Curr_tmStp:(Curr_tmStp+1) , ])), ID=Curr_tmStp) )) # #Conver line to spatial object for plot
    plot(StepAsLine,add=T, col= Color_indv[Curr_indv],lwd=4  ) #plot the buffer
  }#plot only up to 10 indiv
  
  }#loop on invdividuals
  print(c("done with timestep",Curr_tmStp,"out of", N_tmStp))
      
}#loop on time steps
      
      
#### after the nested loops -how many times out of the box? ####
startIndx=1;
for (k in 1:N_indv) {        
  endIndx=startIndx+N_tmStp-1;
  XYind_log2[c(startIndx:endIndx),c("x","y")]=XYind[[k]]        
  XYind_log2[c(startIndx:endIndx),c("pseudoSex")]=  HRCnt[k,3] #storing the sex of this agent sbased on the HR center it got was it a male1 or a female2 there?
  startIndx=endIndx+1;
}        
OutOfTheBox=sum((XYind_log2$x<  -Scl/2) | 
                  (XYind_log2$x>   Scl/2) |
                  (XYind_log2$y<  -Scl/2) |
                  (XYind_log2$y>   Scl/2) ) /length(XYind_log2$x)
print(paste("out of the box rate",round(OutOfTheBox,digit=5)))
rm("Curr_indv","Curr_tmStp","k","Color_indv","PointsStrc","CurDrift","BiasPoint")
rm("Phi_ind","step","step.len","mu","mu.av","coo","StepAsLine","Dist", "endIndx","startIndx","ToPlot")


###### Saving results for matlab and R#####
Name1=paste("xyFromSimulationForSNanalysis", N_tmStp , N_indv,  100*EtaCRW, StpSize_ind,DriftHRCenters,".rdata",sep="_")
Name2="xyFromSimulationForSNanalysis.mat"
save(list=ls(),file=Name1)
#writeMat(Name2,XY=XYind_log2,HRCntXY=HRCnt) 
print('saved simulation data, go to matlab to run the nalysis... ')



```

